Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-03 paper/paper.tex
included keywords
fix fig 1
fix figs 6, 7 and 8
pass

@@

%% \begin{IEEEkeywords}
%% component, formatting, style, styling, insert
%% \end{IEEEkeywords}

@@

\begin{wrapfigure}[7]{r}{0.21\textwidth}
    \vspace{-4ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=3pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            ymax=250,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

First, it analyzes the malicious traffic to extract rule's action,
header, and \emph{initial} options expressed in the
traffic. Considering the Ping Scan attack, \tname{} replays the
attack---with NMAP, in this case---and monitors the traffic.  It
infers the header by checking the protocol of the input messages. In
this case, \tname{} sets the header to \CodeIn{icmp any any -> any
  any}. It uses the most general flow description to command the
\nids\ to monitor the icmp traffic flowing from/to any address and
port.  The rationale is that the choice of these parameters are
domain-specific and subject to change by sys admins. Likewise,
\tname{} leaves the action of the corresponding rule constant, as
\CodeIn{alert}. Furthermore, recall from
Section~\ref{sec:rules-and-packets} that some options can be inferred
from packet contents. \tname{} finds the following five options at
this stage: {\scriptsize{\texttt{\textbf{dsize}:0; \textbf{itype:}8;
      \textbf{icode:}0; \textbf{icmp\_id:}23570;
      \textbf{icmp\_seq:}3439;}}}.  By construction, a rule that uses
this set of options captures the malicious traffic. Unfortunately, the
rule above is over-specified. Informally, the rule includes ``too
specific'' characteristics of the malicious traffic and could
potentially result in false negatives. In this case, although the
options \CodeIn{dsize:0} and \CodeIn{itype:8} are present in the set
of inferred options---as the golden rule from
Figure~\ref{fig:pingscan-example} shows---the others are not.

@@

Second, \tname{} uses the benign traffic to mitigate the over-fitting
problem. In this step, \tname{} searches for alternative rules that
preserve the invariant to capture the positive (/malicious) traffic
but not capture the negative benign traffic. It uses the rule obtained
in the previous step as seed to bootstrap the search for alternative
plausible rules, \ie{}, rules that satisfy the aforementioned
invariant. The search is designed to generate rules containing subsets
of the options declared in the seed rule. Considering this example,
\tname{} produces a total of \pingscanplausible{} plausible rules at
the end of the search. This is a smaller number when compared to $2^n$
option combinations, where $n$ is the number of options contained in
the seed rule ($n=5$, in this case). \tname{} is able to prune the
search space and discard many implausible rules. It is worth noting
that the number of plausible rules that \tname\ produces at this step
can be very high for other example attacks. For illustration, see
column ``\#~Plausible Rules'' on Table~\ref{table:results}.

@@

Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
payload.  Figure~\ref{fig:distribution-contents} shows the
distribution of number of content options per rule for the data set
mentioned above. We limited the number of content options per rule to
10 in the histogram for space. \MyComment{Rules without
  \CodeIn{content} options are an exception and half of the rules
  contain at least two \CodeIn{content} options.}  To sum up, handling
content options for http is very important.

@@

Line~\ref{line-init} initializes the current population of individuals
(\ie{}, rules) $\mathcal{R}$ with the seed rule $sr$. We encode a rule
as a bitvector where each index in the vector represents one option;
the value assigned at a given position indicates whether or not the
option is present in the rule. One iteration of the outer loop
(lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
population of individuals $\mathcal R$ and creates the next generation
$\mathcal R'$. Line~\ref{init-rprime} initializes the population of
individuals for the next generation. The inner loop
(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
candidate rule in $\mathcal R$. The algorithm mutates each of these
rules $N$ times, discarding one rule option each
time. Higher (resp., lower) values of $N$ produces more (resp.,
less) rules. We found empirically that a value of 2 was sufficient to
generate rules in all cases. Line~\ref{copy} assigns a copy of the bitvector encoding
$\mathit{ro}$ to variable $r$. Function $\mathit{len}$ returns the
length of a vector and function $\mathit{rand(lo, hi)}$ randomly
chooses one integer value in the range $[lo, hi]$ defined by its
parameters. So, the expression $\mathit{rand(0, len(r))}$, appearing
at line~\ref{choose-index}, returns an index of the option in $r$ and
the statement at line~\ref{delete-option} deletes that option. The
call to $\mathit{isNew}$ checks if $r$ has not been previously
generated in another iteration whereas the call to
$\mathit{isPlausible}$ checks if the rule only captures positive
traffic, \ie{}, it checks that no message present in $\mathit{bTr}$ is
captured by $r$. If $r$ satisfies both criteria, it is added to
$\mathcal R'$, which denotes the next generation of individuals in the
search. After processing every candidate rule in $\mathcal R$, the
next generation of individuals is defined (line~\ref{define-next-gen})
and the set of discovered rules is updated (at
line~\ref{output-accum}).

@@

\textbf{Correctness argument.}~The algorithm reports rules that
capture only the positive traffic. Every rule created during the
search captures the attack associated with the input as every rule
created contains a proper subset of the options from the original rule
and all those options are, by definition, satisfied by the traffic
produced by the original attack. In addition, from the definition of
$\mathit{isPlausible}$, reported rules capture \emph{only} the
positive traffic. If negative traffic is captured, a rule is not added
to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
O$ and reported on the output. Finally, from the definition of
function $\mathit{filter}$, only the rules in $\mathcal O$ that
capture all the instances of the attack are reported on output.
\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
our setup calls \suri\ directly through an API. \suri\ is invoked
multiple times for creating rules.

@@

\tname{} use the popularity of the options used in trusted rules as a
heuristic function---rules containing popular options score higher
compared to those containing unpopular
rules. Figure~\ref{fig:distribution-options} shows the histogram of
rule options for the ruleset we used~\cite{emerging-threats-open}. The
x-axis shows the options used in rules (the 10 most common) and the
y-axis shows the number of rules using the corresponding option. Note
that the options \CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce}
are very popular. \tname\ uses the following formula to compute the
contribution of option popularity to the score of a rule $r$:
\indent
$\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N$.

@@

          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype},          

@@

        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155)};

@@

\vspace{-4ex}

@@

\noindent
where $\mathit{content_i}$ denotes the $i$th content in a given
rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our ruleset. To illustrate, consider a
rule whose content options use the strings ``GET'' and ``ID''. The
contribution of that rule for the computation of similarity score
would be $0.601$ $(=(1$+$.203)$/$2)$. The relatively high weight
reflects the fact that ``GET'' and ``ID'' are popular content strings.

\begin{wrapfigure}[9]{r}{0.21\textwidth}
  \vspace{-2ex}
\centering
\includegraphics[scale=0.6]{figs/http_header-example.png}
\vspace{-1ex}
\caption{Content modifiers.}
\label{fig:http-header-example}
\end{wrapfigure}

@@

\noindent
where $g(i)$ is the contribution of content option $i$ to the score,
$N$ is the number of contents options present in a rule,
$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
of a modifier $j$ to content option $i$,
$\mathit{max(modifier\_frequency)}$ refers to the modifier with
maximum popularity (for normalization), and $M_i$ is the number of
content modifiers that could access the string associated with content
$i$.
To illustrate, consider a rule containing the option
\CodeIn{content:``/HtmlAdaptor''} (see Section~\ref{sec:jboss}). The
contribution of this option (\ie{}, the value of $\mathit{g(i)}$) to
the score of the rule is $0.5$ (=$\frac{10,111+11}{10,111*2}$). The
value of $M_i$, in this case, is 2 as the string ``/HtmlAdaptor''
could be accessed through the modifiers \CodeIn{uri} and
\CodeIn{request\_line}. The reason is that \tname\ identifies that
that string appears in the field \CodeIn{Request URI} (of an http
message from the malicious traffic) and that field is accessible
through these two modifiers. The relatively high heuristic value
reflects the fact that the string could be accessed by two popular
content modifiers.

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule in the ranking for an increasing number
\begin{wrapfigure}[9]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-2ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{wrapfigure}[8]{r}{0.28\textwidth}  
  \vspace{-2.5ex}
  \centering
  %% \begin{subfigure}{.2\textwidth}
  %%   \centering
  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
  %%   \vspace{-1ex}        
  %%   \caption{coldfusion}
  %% \end{subfigure}%
  %% \hspace{-8ex}
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-3ex}    
    \caption{inc}    
  \end{subfigure}%
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-3ex}    
    \caption{jsp}    
  \end{subfigure}%
  \vspace{-1ex}
  \caption{Negative examples.}
  \vspace{-2ex}
  \label{fig:impact-negative}
\end{wrapfigure}

>>>>

\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

@@

\keywords{Intrusion Detection, Suricata, Rule-based, Synthesis}

% \begin{IEEEkeywords}
%% component, formatting, style, styling, insert
%% \end{IEEEkeywords}

@@

\begin{wrapfigure}[8]{r}{0.21\textwidth}
    \vspace{-4ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=3pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            ymax=250,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of Snort and \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

First, it analyzes the malicious traffic to extract rule's action,
header, and \emph{initial} options expressed in the
traffic. Considering the Ping Scan attack, \tname{} 
infers the header by checking the protocol of the input messages. In
this case, \tname{} sets the header to \CodeIn{icmp any any -> any
  any}. It uses the most general flow description to command the
\nids\ to monitor the icmp traffic flowing from/to any address and
port.  The rationale is that the choice of these parameters are
domain-specific and subject to change by sys admins. Likewise,
\tname{} leaves the action of the corresponding rule constant, as
\CodeIn{alert}. Furthermore, recall from
Section~\ref{sec:rules-and-packets} that some options can be inferred
from packet contents. \tname{} finds the following five options at
this stage: {\scriptsize{\texttt{\textbf{dsize}:0; \textbf{itype:}8;
      \textbf{icode:}0; \textbf{icmp\_id:}23570;
      \textbf{icmp\_seq:}3439;}}}.  By construction, a rule that uses
this set of options captures the malicious traffic. Unfortunately, the
rule above is over-specified. Informally, the rule includes ``too
specific'' characteristics of the malicious traffic and could
potentially result in false negatives. In this case, although the
options \CodeIn{dsize:0} and \CodeIn{itype:8} are present in the set
of inferred options---as the golden rule from
Figure~\ref{fig:pingscan-example} shows---the others are not.

@@

Second, \tname{} uses the benign traffic to mitigate the over-fitting
problem. In this step, \tname{} searches for alternative rules that
preserve the invariant to capture the positive (malicious) traffic
but not capture the negative (benign) traffic. It uses the rule obtained
in the previous step as seed to bootstrap the search for alternative
plausible rules, \ie{}, rules that satisfy the aforementioned
invariant. The search is designed to generate rules containing subsets
of the options declared in the seed rule. Considering this example,
\tname{} produces a total of \pingscanplausible{} plausible rules at
the end of the search. This is a smaller number when compared to $2^n$
option combinations, where $n$ is the number of options contained in
the seed rule ($n=5$, in this case). \tname{} is able to prune the
search space and discard many implausible rules. It is worth noting
that the number of plausible rules that \tname\ produces at this step
can be very high for other example attacks. For illustration, see
column ``\#~Plausible Rules'' on Table~\ref{table:results}.

@@

Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
payload.  Figure~\ref{fig:distribution-contents} shows the
distribution of number of content options per rule for the data set
mentioned above. We limited the number of content options per rule to
8 in the histogram for space. \MyComment{Rules without
  \CodeIn{content} options are an exception and half of the rules
  contain at least two \CodeIn{content} options.}  To sum up, handling
content options for http is very important.

@@

Line~\ref{line-init} initializes the current population of individuals
(\ie{}, rules) $\mathcal{R}$ with the seed rule $sr$. We encode a rule
as a bitvector where each index in the vector represents one option;
the value assigned at a given position indicates whether or not the
option is present in the rule. One iteration of the outer loop
(lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
population of individuals $\mathcal R$ and creates the next generation
$\mathcal R'$. Line~\ref{init-rprime} initializes the population of
individuals for the next generation. The inner loop
(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
candidate rule in $\mathcal R$. The algorithm mutates each of these
rules $N$ times, discarding one rule option each
time. Higher (respectively, lower) values of $N$ produces more (respectively,
less) rules. We found empirically that a value of 2 was sufficient to
generate the golden rule in all cases. Line~\ref{copy} assigns a copy of the bitvector encoding
$\mathit{ro}$ to variable $r$. Function $\mathit{len}$ returns the
length of a vector and function $\mathit{rand(lo, hi)}$ randomly
chooses one integer value in the range $[lo, hi]$ defined by its
parameters. So, the expression $\mathit{rand(0, len(r))}$, appearing
at line~\ref{choose-index}, returns an index of the option in $r$ and
the statement at line~\ref{delete-option} deletes that option. The
call to $\mathit{isNew}$ checks if $r$ has not been previously
generated in another iteration whereas the call to
$\mathit{isPlausible}$ checks if the rule only captures positive
traffic, \ie{}, it checks that no message present in $\mathit{bTr}$ is
captured by $r$. If $r$ satisfies both criteria, it is added to
$\mathcal R'$, which denotes the next generation of individuals in the
search. After processing every candidate rule in $\mathcal R$, the
next generation of individuals is defined (line~\ref{define-next-gen})
and the set of discovered rules is updated (at
line~\ref{output-accum}).

@@

\textbf{Correctness argument.}~The algorithm reports rules that
capture only the positive traffic. Every rule created during the
search captures the attack associated with the input as every rule
created contains a proper subset of the options from the original rule
and all those options are, by definition, satisfied by the traffic
produced by the original attack. In addition, from the definition of
$\mathit{isPlausible}$, reported rules capture \emph{only} the
positive traffic. If negative traffic is captured, a rule is not added
to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
O$ and reported on the output. Finally, from the definition of
function $\mathit{filter}$, only the rules in $\mathcal O$ that
capture all the instances of the attack are reported on output.
\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
our setup calls \suri\ offiline by providing a \emph{pcap} file as input. \suri\ is invoked
multiple times for creating rules.

@@

\tname{} use the popularity of the options used in trusted rules as a
heuristic function---rules containing popular options score higher
compared to those containing unpopular
options. Figure~\ref{fig:distribution-options} shows the histogram of
rule options for the ruleset we used~\cite{emerging-threats-open}. The
x-axis shows the options used in rules (the 8 most common) and the
y-axis shows the number of rules using the corresponding option. Note
that the options \CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce}
are very popular. \tname\ uses the following formula to compute the
contribution of option popularity to the score of a rule $r$:
\indent
$\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N$.

@@

          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode},          

@@

        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode,108)};

@@

@@

\noindent
where $\mathit{content_i}$ denotes the $i$th content in a given
rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the seven most popular strings in our ruleset. To illustrate, consider a
rule whose content options use the strings ``GET'' and ``id=''. The
contribution of that rule for the computation of similarity score
would be $0.601$ $(=(1$+$.203)$/$2)$. The relatively high weight
reflects the fact that ``GET'' and ``id='' are popular content strings.

\begin{wrapfigure}[10]{r}{0.21\textwidth}
  \vspace{-2ex}
\centering
\includegraphics[scale=0.6]{figs/http_header-example.png}
\vspace{-1ex}
\caption{Content modifiers.}
\label{fig:http-header-example}
\end{wrapfigure}

@@

\noindent
where $g(i)$ is the contribution of content option $i$ to the score,
$N$ is the number of contents options present in a rule,
$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
of a modifier $j$ to content option $i$,
$\mathit{max(modifier\_frequency)}$ refers to the modifier with
maximum popularity (for normalization), and $M_i$ is the number of
content modifiers that could access the string associated with content
$i$.
To illustrate, consider a rule containing the option
\CodeIn{content:``/HtmlAdaptor''} (see Section~\ref{sec:jboss}). The
contribution of this option (\ie{}, the value of $\mathit{g(i)}$) to
the score of the rule is $0.5$ (=$\frac{10,111+11}{10,111*2}$). The
value of $M_i$, in this case, is 2 as the string ``/HtmlAdaptor''
could be accessed through the modifiers \CodeIn{uri} and
\CodeIn{request\_line}. The reason is that \tname\ identifies that
that string appears in the field \CodeIn{Request URI} (of an http
message from the malicious traffic) and that field is accessible
through these two modifiers. The relatively high heuristic value
reflects the fact that the string could be accessed by a popular
content modifier.

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule in the ranking for an increasing number
\begin{wrapfigure}[11]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-2ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{wrapfigure}[10]{r}{0.28\textwidth}  
  \vspace{-2.5ex}
  \centering
  %% \begin{subfigure}{.2\textwidth}
  %%   \centering
  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
  %%   \vspace{-1ex}        
  %%   \caption{coldfusion}
  %% \end{subfigure}%
  %% \hspace{-8ex}
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-3ex}    
    \caption{inc}    
  \end{subfigure}%
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-3ex}    
    \caption{jsp}    
  \end{subfigure}%
  \vspace{-1ex}
  \caption{Negative examples.}
  \vspace{-2ex}
  \label{fig:impact-negative}
\end{wrapfigure}
